---
title: "ASU Designation Program"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
runtime: shiny
---

```{r setup, include=FALSE}
library(shiny)
library(flexdashboard)
library(mapgl)
library(tigris)
library(sf)
library(sfdep)
library(tidyverse)
library(readxl)

# ---- tigris cache -------------------------------------------------
# put cached ZIPs in  inst/shiny_app/_tigris_cache  (any writable path works)
tigris::tigris_cache_dir( here::here("inst", "shiny_app", "_tigris_cache") )

# tell tigris to look there first before re‑downloading
options(tigris_use_cache = TRUE)

# ---- resolve paths from the project root ----------------------------
# 'here()' always points at the directory that contains ASUbuildR.Rproj
# no matter where this .Rmd lives.
if (!requireNamespace("here", quietly = TRUE))
  install.packages("here")

library(here)          # <‑‑ add once
source(here::here("R", "run_asu_original.r"))   # local = FALSE by default
source(here::here("R", "run_tract_hunter.R"))
options(shiny.maxRequestSize = 50 * 1024^2)  # Set to 50MB or your desired limit

# Initialize reactive values
uploaded_data <- reactiveVal(NULL)
state <- reactiveVal(NULL)
tract_year <- reactiveVal(NULL)
tract_data <- reactiveVal(NULL)
asu_data <- reactiveVal(NULL)
asu_tracts <- reactiveVal(NULL)
asu_summary <- reactiveVal(NULL)
full_data <- reactiveVal(NULL)
full_data_reset <- reactiveVal()
selected_tracts <- reactiveVal(NULL)
edit_table <- reactiveVal(NULL)
edit_summary <- reactiveVal(NULL)
current_selection <- reactiveVal(NULL)
map_data_dbg <- reactiveVal(NULL)   # container




# Load static data
excel_names <- c("record",
                 "geoid",
                 "st_fips",
                 "cnty_fips",
                 "tract_fips",
                 "name",
                 "tract_pop_dec",
                 "tract_pop_cur",
                 "tract_emp",
                 "tract_unemp",
                 "tract_urate",
                 "tract_urate_error",
                 "cnty_pop_dec",
                 "cnty_pop_cur",
                 "cnty_emp",
                 "cnty_unemp",
                 "cnty_urate",
                 "cnty_urate_error",
                 "pop_shr",
                 "emp_shr",
                 "unemp_shr",
                 "laus_primary",
                 "cnty_emp_ave",
                 "cnty_unemp_ave",
                 "tract_ASU_clf",
                 "tract_ASU_emp",
                 "tract_ASU_unemp",
                 "tract_ASU_urate")


```

Data Initialization
=============================

## Controls {data-width=150}

This screen allows you to import the BLS-provided ASU file, generally named like *NV_asu23.xlsx* which will be the only external data import step necessary.  The script will read the file, extract the state and corresponding year, and standardize the column names to avoid annual updates to the script.

If the necessary data is not in columns A through AB within the Excel file, this process will not work correctly.

Once the file is loaded, a preview of the data will appear on the right, the state and year will display below, and the analyst may proceed to the next tab (Load Initial ASU).

```{r}
fileInput("file", "Choose Excel File", accept = c(".xls", ".xlsx"))

textOutput("selected_state")

textOutput("selected_year")


```

## Output {data-width=850}

```{r}

tableOutput("data_preview")

observeEvent(input$file, {
  req(input$file)
  df <- read_excel(input$file$datapath, range = "A2:AB25000", col_names = excel_names) %>%
    filter(!is.na(geoid)) %>%
    mutate(GEOID = str_remove(geoid, "14000US")) %>%
    select(GEOID,
           st_fips,
           cnty_fips,
           tract_fips,
           name,
           tract_pop_cur,
           tract_ASU_clf,
           tract_ASU_emp,
           tract_ASU_unemp,
           tract_ASU_urate)
  
  uploaded_data(df)
  
  uploaded_data() %>%
    pull(st_fips) %>%
    unique() %>%
    state()
  
  read_excel(input$file$datapath, range = "H1:H1", col_names = "year") %>%
    pull(year) %>%
    str_remove("tract_pop") %>%
    tract_year()
  
  output$data_preview <- renderTable({
    head(uploaded_data(),50)
    
  })
  
  output$selected_state <- renderText({
    paste0("Selected FIPS is: ", state())
  })
  
    output$selected_year <- renderText({
    paste0("Tract Population Year is: ", tract_year())
  })
  
})


```

Load Initial ASU
============================

## Controls {data-width=200}

This button will launch two processes, which will take up to 5 minutes to complete. For smaller states, it may take under 15 seconds.

First, the application will download census shapefiles by tract for the appropriate state and year. Next, after combining the BLS file with these shapefiles, it will iteratively search for contiguous geographic regions until the combined unemployment rate in each region is as close to 6.45% as possible.

When this process is complete, a map will appear to the right, and a summary of the created ASU regions will appear below.  It is important to note, these will still include areas that do not qualify as an independent ASU (there is no check for population), but it gives the analyst a starting place for analysis.

Once this step is done, you may proceed to the next step.

```{r}
radioButtons(
  "asu_algo", "Choose algorithm:",
  choices = c(
    "Original" = "orig",
    "Tract Hunter" = "mine"
  ),
  inline = TRUE
)

checkboxInput(
  inputId = "join_touching",
  label   = "Include touching tracts when building ASUs",
  value   = TRUE          # default keeps current behaviour
)


actionButton("process", "Load Tracts and Initialize ASU")
tableOutput("asu")

observeEvent(input$process, {

  ## 1 ─ Build the base tract + BLS data ------------------------------
  tract_list <- tracts(state = state(), year = tract_year()) |>
    mutate(
      row_num    = row_number(),
      continuous = st_contiguity(geometry)
    )

  data_merge <- tract_list |>
    left_join(uploaded_data(), by = "GEOID")
    # arrange(desc(tract_ASU_urate), desc(tract_ASU_unemp))  # not needed here

  ## 2 ─ Run the chosen ASU algorithm ---------------------------------
  algo_choice <- isolate(input$asu_algo %||% "orig")
  join_flag   <- isTRUE(input$join_touching)         # << the new toggle

  res <- if (algo_choice == "orig") {
           run_asu_original(tract_list, uploaded_data())
         } else {
           run_tract_hunter(tract_list, uploaded_data(),
                            join_touching = join_flag)
         }

  ## 3 ─ Ensure MULTIPOLYGON for maplibregl ---------------------------
  if (!inherits(res$full_data, "sf"))
    stop("res$full_data is not an sf object")

  res$full_data <- sf::st_cast(res$full_data, "MULTIPOLYGON", warn = FALSE)

  ## 4 ─ Push results into reactiveVals already used elsewhere --------
  full_data(      res$full_data)
  full_data_reset(res$full_data_reset)
  asu_data(       res$asu_data)
  asu_tracts(     res$asu_tracts)
  asu_summary(    res$asu_summary)
  map_data_dbg(   res$full_data)

  ## 5 ─ Render the summary table just under the button ---------------
  output$asu <- renderTable({ asu_summary() })

  ## 6 ─ (Re)draw the first map ---------------------------------------
  output$initial_map <- renderMaplibre({

    req(full_data())
    map_data <- full_data()

    # ---- 6·1 geometry housekeeping ----------------------------------
    map_data <- map_data |>
      sf::st_make_valid() |>
      sf::st_zm(what = "ZM", drop = TRUE) |>
      dplyr::filter(!sf::st_is_empty(geometry)) |>
      dplyr::mutate(geom_type = sf::st_geometry_type(geometry)) |>
      dplyr::filter(geom_type %in% c("POLYGON", "MULTIPOLYGON")) |>
      dplyr::select(-geom_type)

    if (nrow(map_data) == 0) {
      return(
        maplibre(style = carto_style("positron")) |>
        add_control("No valid polygon geometry to display.", "top-left")
      )
    }

    # ---- 6·2 palette / breaks ---------------------------------------
    map_data <- map_data |>
      dplyr::mutate(asunum = dplyr::na_if(asunum, 0L))

    n_asu <- max(map_data$asunum, na.rm = TRUE)
    n_asu <- ifelse(is.finite(n_asu), n_asu, 0L)

    palette_logic <- function(n_asu) {
      if (n_asu == 0) {
        list(fill = "#808080", legend_vals = "(none)", legend_cols = "#808080")
      } else if (n_asu == 1) {
        list(
          fill = step_expr(
            column   = "asunum",
            base     = "#808080",
            stops    = "#7FC97F",
            values   = 1,
            na_color = "#444444"
          ),
          legend_vals = 1,
          legend_cols = "#7FC97F"
        )
      } else {
        n_cat  <- min(max(n_asu, 3L), 8L)
        pal    <- RColorBrewer::brewer.pal(n_cat, "Accent")
        breaks <- seq(1 + n_asu / n_cat, n_asu, length.out = n_cat - 1)
        list(
          fill = step_expr(
            column   = "asunum",
            base     = pal[1],
            stops    = pal[-1],
            values   = breaks,
            na_color = "#444444"
          ),
          legend_vals = seq_len(n_asu),
          legend_cols = pal
        )
      }
    }

    pal <- palette_logic(n_asu)

    # ---- 6·3 build map ----------------------------------------------
    maplibre(style = carto_style("positron")) |>
      fit_bounds(map_data, animate = FALSE) |>
      add_fill_layer(
        id                 = "basemap",
        source             = map_data,
        fill_color         = pal$fill,
        fill_opacity       = 0.5,
        fill_outline_color = "black",
        popup = concat(
          "<strong>ASU Number: </strong>", get_column("asunum"), "<br>",
          "<strong>Tract GEOID: </strong>", get_column("GEOID"),  "<br>",
          "<strong>UR: </strong>",
          number_format(get_column("tract_ASU_urate"),
                        style = "decimal", maximum_fraction_digits = 2),
          "%<br>",
          "<strong>Unemp: </strong>",
          number_format(get_column("tract_ASU_unemp"),
                        style = "decimal", maximum_fraction_digits = 0), "<br>"
        )
      ) |>
      add_legend(
        "ASU Number",
        values = pal$legend_vals,
        colors = pal$legend_cols,
        type   = "categorical"
      )
  })

})  # ---- end observeEvent -------------------------------------------------


# ─────────────── DEBUG PANEL ───────────────
output$map_data_preview <- renderPrint({
  # show something only after the ASU routine ran
  req(map_data_dbg())

  # ▸ just a quick glimpse – change to str() or summary() if you prefer
  cat("----- head(map_data_dbg()) -----\n")
  print(
    map_data_dbg() |>
      sf::st_drop_geometry() |>      # geometry is noisy in a console print
      dplyr::slice_head(n = 10)
  )
  cat("\nRows :", nrow(map_data_dbg()),
      "   Cols :", ncol(map_data_dbg()), "\n")
})

```

## Output {data-width=800}

```{r}
maplibreOutput("initial_map", height = "100vh")

```

Modify ASU Selections
==========================

## Controls {data-width=150}

Selecting tracts on the map to the right will allow you to modify the designated ASU from the inital selection methodology. Several tools are provided to manage this process.

You may select tracts by clicking on the map, or by selecting a numbered ASU from the dropdown below to select all tracts in that ASU.  The *Preview* button will calculate metrics for all selected tracts.

You may also choose to *Filter* selected tracts to find the lowest unemployment rates in an area.  This can be useful to omit low-unemployment tracts after adding tracts to connect isolated regions.

To modify the assigned ASU, with tracts selected, use the *Update* button to update the ASU of selected tracts. To remove a tract from any ASUs, assign it to ASU zero.

To undo your work and return to the initially-designated data, use the *Reset* button.

```{r}
# Select an ASU
asu_selection_ui <- reactive({
  selectInput("select_asunum", "Select ASU Number:",
              choices = c("None", asu_choices()),
              selected = "None")
})
renderUI(asu_selection_ui())
renderUI(HTML("<hr>"))

# Change ASU
numericInput("new_asu", "Change ASU Value:", value = 0, min = 0)
actionButton("update", "Update Selected Tracts")
renderUI(HTML("<hr>"))

# Clear Selected Tracts
actionButton("clear", "Clear Selected Tracts")
renderUI(HTML("<hr>"))

# Show Data for Selection
actionButton("preview", "Show Data for Selected Tracts")
asu_choices <- reactive({
  req(full_data())                     # only run after full_data exists
  full_data() |> pull(asunum) |> unique() |> sort()
})


# Show Lowest UR Tracts in Selection
numericInput("percentile", "Percentile Threshold:", value = 5, min = 1, max = 100, step = 1)
actionButton("filter_percentile", "Show Tracts with Lowest UR")
renderUI(HTML("<hr>"))


# Single GROID Selection
textInput("manual_geoid", "Enter a GEOID:", placeholder = "32031990000")
actionButton("select_geoid", "Select a Single Tract")
renderUI(HTML("<hr>"))

# Reset to Initial Load
actionButton("reset", "Reset to Initial Data")



### Need to add new inputs / outputs
# Add table of currently selected tracts
# ensure basemap includes all tracts, not just those with an ASU assigned


```

## Output {data-width=850}

```{r}

# 
# edit_table <- reactiveVal(data.frame(GEOID = character(),
#                                      tract_pop_cur = numeric(),
#                                      tract_ASU_clf = numeric(),
#                                      tract_ASU_unemp = numeric(),
#                                      tract_ASU_urate = numeric(),
#                                      stringsAsFactors = FALSE))

maplibreOutput("edit_map", height = "90vh")

tableOutput("selection_data")

tableOutput("selection_summary")



output$edit_map <- renderMaplibre({
  
  req(full_data())
  
  number_asus <- max(full_data()$asunum)
  number_categories <- min(c(number_asus, 8))
  asu_per_category <- number_asus/number_categories
  
  # Create breaks for ASU categories, excluding 0
  if (number_categories == 1) {
    # Single ASU: just need one break at 1
    asu_breaks <- 1
    actual_categories <- 1
    brewer_pal <- "#7FC97F"  # Single color
  } else {
    # Multiple ASUs: create breaks from 1 to max, evenly spaced
    asu_breaks <- seq(from = 1, to = number_asus, length.out = number_categories)
    actual_categories <- max(3, number_categories)  # Ensure minimum of 3 for brewer.pal
    full_brewer_pal <- RColorBrewer::brewer.pal(actual_categories, "Accent")
    brewer_pal <- full_brewer_pal[1:number_categories]  # Take only what we need
  }

  maplibre(style = carto_style("positron")) |>
    fit_bounds(full_data(), animate = FALSE) |> 
    add_fill_layer(
      id = "basemap",
      source = full_data(),
      fill_color = step_expr(
        column = "asunum",
        base = "#808080",  # Grey for asunum == 0 (and anything < first break)
        stops = brewer_pal,
        values = asu_breaks,
        na_color = "#444444"
      ),
      fill_opacity = 0.5,
      fill_outline_color = "black",
      tooltip = concat(
        "<strong>ASU Number: </strong>", get_column("asunum"), "<br>",
        "<strong>Tract GEOID: </strong>", get_column("GEOID"), "<br>",
        "<strong>Tract Unemployment Rate: </strong>", number_format(get_column("tract_ASU_urate"), style = "decimal", maximum_fraction_digits = 2), "%<br>",
        "<strong>Tract Unemployment: </strong>", number_format(get_column("tract_ASU_unemp"), style = "decimal", maximum_fraction_digits = 0), "<br>"
      )
    )
})

# Corrected click event handler for MapLibre
observeEvent(input$edit_map_feature_click, {
  click <- input$edit_map_feature_click
  current_selection <- selected_tracts()
  
  # Extract the GEOID from the clicked feature properties
  clicked_geoid <- click$properties$GEOID
  
  # Check if clicked_geoid is not NULL and not empty
  if (is.null(clicked_geoid) || length(clicked_geoid) == 0 || is.na(clicked_geoid)) {
    return()  # Exit early if no valid GEOID
  }
  
  # Add or remove the selected tract
  if (clicked_geoid %in% current_selection) {
    current_selection <- setdiff(current_selection, clicked_geoid)
  } else {
    current_selection <- c(current_selection, clicked_geoid)
  }
  
  selected_tracts(current_selection)
  
  # Update the map highlighting
  if (length(current_selection) > 0) {
    selected_data <- subset(full_data(), GEOID %in% selected_tracts())
    
    maplibre_proxy("edit_map") %>%
      clear_layer("selected") %>%
      add_line_layer(
        id = "selected",
        source = selected_data,
        line_color = "red",
        line_width = 3,
        line_opacity = 1
      )
  } else {
    # Clear the selection layer if no tracts are selected
    maplibre_proxy("edit_map") %>%
      clear_layer("selected")
  }
  
  # Update the selection table
  output$selection_table <- renderTable({
    selected_tracts()
  })
})

# Clear Selected Tracts on Button Click
observeEvent(input$clear, {
  
  selected_tracts(NULL)
  
  # Update the map highlighting
    maplibre_proxy("edit_map") %>%
      clear_layer("selected")
    return()
})

# Select tracts in an ASU from dropdown
observeEvent(input$select_asunum, {
  req(full_data())
  
  # Skip if "None" is selected
  if (input$select_asunum == "None") {
    selected_tracts(NULL)
    maplibre_proxy("edit_map") %>%
      clear_layer("selected")
    return()
  }
  
  # Get all tracts with the selected ASU number
  selected_tract_ids <- full_data() %>%
    filter(asunum == as.numeric(input$select_asunum)) %>%
    pull(GEOID)
  
  # Update the selected_tracts reactive value
  selected_tracts(selected_tract_ids)
  
  # Get the filtered data for selected tracts
  selected_data <- subset(full_data(), GEOID %in% selected_tract_ids)
  
  # Highlight the selected tracts on the map using maplibre proxy
  maplibre_proxy("edit_map") %>%
    clear_layer("selected") %>%
    add_line_layer(
      id = "selected",
      source = selected_data,
      line_color = "red",
      line_width = 3,
      line_opacity = 1
    ) |> 
    fit_bounds(selected_data, animate = TRUE)
})

# Select a single tract given a GEOID
observeEvent(input$select_geoid, {
  req(full_data())
  
  # Get the selected tract's data
  selected_tract <- full_data() %>%
    filter(GEOID == input$manual_geoid)
  
  # If the tract exists, highlight it and fit bounds
  if (nrow(selected_tract) > 0) {
    # Highlight the selected tract on the map using maplibre proxy
    maplibre_proxy("edit_map") %>%
      clear_layer("selected") %>%
      add_line_layer(
        id = "selected",
        source = selected_tract,
        line_color = "red",
        line_width = 3,
        line_opacity = 1
      ) %>%
      fit_bounds(selected_tract, animate = TRUE)
    
    # Update the selected_tracts reactive value
    selected_tracts(selected_tract$GEOID)
  }
})



# Filter selected tracts by percentile
observeEvent(input$filter_percentile, {
  
  req(input$percentile, full_data(), selected_tracts())
  
  # Calculate the threshold for the percentile based on selected tracts
  selected_data <- full_data() %>%
    filter(GEOID %in% selected_tracts())
  
  threshold <- quantile(selected_data$tract_ASU_urate, input$percentile/100)
  
  # Filter currently selected tracts by the new percentile threshold
  filtered_selected <- selected_data %>%
    filter(tract_ASU_urate <= threshold,
           tract_ASU_urate > 0) %>%
    pull(GEOID)
  
  # Update the selected_tracts reactive value with filtered selection
  selected_tracts(filtered_selected)
  
  # Get the filtered data for selected tracts
  filtered_data <- subset(full_data(), GEOID %in% filtered_selected)
  
  # Update the map to highlight the newly filtered tracts using maplibre proxy
  maplibre_proxy("edit_map") %>%
    clear_layer("selected") %>%
    add_line_layer(
      id = "selected",
      source = filtered_data,
      line_color = "red",
      line_width = 3,
      line_opacity = 1
    )
  
  # Update the selection table
  output$selection_table <- renderTable({
    filtered_selected
  })
  
})


# Calculate and display a table of data for selected tracts
observeEvent(input$preview, {
  req(selected_tracts())

  replacement_data <- full_data() %>%
    st_drop_geometry() %>%
    subset(GEOID %in% selected_tracts())

  edit_table(replacement_data)

  edit_table() %>%
    ungroup() %>%
    summarize(
      tracts = n(),
      population = sum(tract_pop_cur, na.rm = TRUE),
      lf = sum(tract_ASU_clf, na.rm = TRUE),
      unemp = sum(tract_ASU_unemp, na.rm = TRUE),
      ur = round(unemp/lf*100, 5)
    ) %>%
    edit_summary()

  # Show Preview of Selected Tracts
  output$selection_summary <- renderTable({
    edit_summary()
  })

})

# Update selected tracts to new ASU number.
observeEvent(input$update, {
  req(selected_tracts(), full_data())
  
  # Update the full_data reactive value
  updated_data <- full_data() %>%
    filter(GEOID %in% selected_tracts()) %>%
    mutate(asunum = as.numeric(input$new_asu)) %>%
    rbind(
      full_data() %>%
        filter(!(GEOID %in% selected_tracts()))
    )
  
  # Update the reactive value
  full_data(updated_data)
  
  # Clear selected tracts
  selected_tracts(NULL)
  
  # Update map to remove highlight on selected tracts
  maplibre_proxy("edit_map") %>%
    clear_layer("selected")
})


# Reset data to initial data load from Excel input
observeEvent(input$reset, {
  req(full_data_reset(), full_data())
  
  # Reset the full_data reactive value
  full_data(full_data_reset())
  
  # Clear selected tracts
  selected_tracts(NULL)
  
  # Update map to remove highlight on selected tracts
  maplibre_proxy("edit_map") %>%
    clear_layer("selected")
})




```

Save and Load Data
==========================

## Controls  {data-width=150}

These buttons allow you to save the map in progress so that you can load it again in the future to continue your work.

First, choose a save directory. The default is your current working directory. After loading data, you can go directly to the ASU Modification tab to continue making edits.

```{r}

# Directory selection - Simple text input approach
textInput("save_dir_path", "Save Directory Path:", value = getwd(), width = "100%")
br()
actionButton("browse_dir", "Browse for Directory")
br()
br()
textOutput("current_save_dir")
textOutput("dir_status")
br()

actionButton("save_data", "Save Data")
fileInput("load_data", "Load Data", accept = c(".rds"))

# Reactive value to store the selected directory
save_directory <- reactiveVal(getwd())  # Default to current working directory

# Observer to handle directory path changes
observeEvent(input$save_dir_path, {
  if (!is.null(input$save_dir_path) && input$save_dir_path != "") {
    if (dir.exists(input$save_dir_path)) {
      save_directory(input$save_dir_path)
    }
  }
})

# Alternative browse button (opens file dialog for directory selection)
observeEvent(input$browse_dir, {
  # This will work in RStudio or desktop R
  if (interactive()) {
    selected_dir <- tryCatch({
      choose.dir(default = save_directory(), caption = "Select Save Directory")
    }, error = function(e) {
      # For non-Windows systems, use tcltk
      tryCatch({
        library(tcltk)
        tk_choose.dir(default = save_directory(), caption = "Select Save Directory")
      }, error = function(e2) {
        NULL
      })
    })
    
    if (!is.null(selected_dir) && selected_dir != "") {
      updateTextInput(session, "save_dir_path", value = selected_dir)
      save_directory(selected_dir)
    }
  } else {
    # In web-deployed apps, show instruction
    showModal(modalDialog(
      title = "Directory Selection",
      "In web-deployed apps, please manually enter the directory path in the text field above.",
      easyClose = TRUE
    ))
  }
})

# Display current save directory
output$current_save_dir <- renderText({
  paste("Current save directory: ", save_directory())
})

# Display directory status
output$dir_status <- renderText({
  if (dir.exists(save_directory())) {
    "✓ Directory exists and is accessible"
  } else {
    "⚠ Directory does not exist - it will be created when saving"
  }
})

# Observer for saving the data
observeEvent(input$save_data, {
  req(full_data())
  
  # Use the selected directory
  save_path <- file.path(save_directory(), "saved_data.rds")
  
  # Check if directory exists, create if it doesn't
  if (!dir.exists(save_directory())) {
    dir.create(save_directory(), recursive = TRUE)
  }
  
  # Save the full_data object to an RDS file
  tryCatch({
    saveRDS(full_data(), save_path)
    
    # Provide feedback to the user
    showModal(modalDialog(
      title = "Data Saved",
      paste("Your data has been saved to:", save_path),
      easyClose = TRUE
    ))
  }, error = function(e) {
    showModal(modalDialog(
      title = "Error Saving Data",
      paste("Failed to save data:", e$message),
      easyClose = TRUE
    ))
  })
})

# Observer for loading the data
observeEvent(input$load_data, {
  req(input$load_data)
  
  # Read the RDS file into a data frame
  loaded_data <- tryCatch({
    readRDS(input$load_data$datapath)
  }, error = function(e) {
    showModal(modalDialog(
      title = "Error Loading Data",
      "Failed to load the data. Please check the file format.",
      easyClose = TRUE
    ))
    return(NULL)
  })
  
  # If loaded_data is NULL (failed to load), exit the observer
  req(!is.null(loaded_data))
  
  # Ensure the loaded data is an sf object
  if (!inherits(loaded_data, "sf")) {
    showModal(modalDialog(
      title = "Error Loading Data",
      "The loaded data is not a valid sf object.",
      easyClose = TRUE
    ))
    return(NULL)
  }
  
  # Update full_data with the loaded data
  full_data(loaded_data)
  
  # Provide feedback to the user
  showModal(modalDialog(
    title = "Data Loaded",
    "Your data has been successfully loaded.",
    easyClose = TRUE
  ))
  
  # Recreate the color scheme logic (same as in renderMaplibre)
  number_asus <- max(full_data()$asunum)
  number_categories <- min(c(number_asus, 8))
  asu_per_category <- number_asus/number_categories
  
  # Create breaks for ASU categories, excluding 0
  if (number_categories == 1) {
    # Single ASU: just need one break at 1
    asu_breaks <- 1
    actual_categories <- 1
    brewer_pal <- "#7FC97F"  # Single color
  } else {
    # Multiple ASUs: create breaks from 1 to max, evenly spaced
    asu_breaks <- seq(from = 1, to = number_asus, length.out = number_categories)
    actual_categories <- max(3, number_categories)  # Ensure minimum of 3 for brewer.pal
    full_brewer_pal <- RColorBrewer::brewer.pal(actual_categories, "Accent")
    brewer_pal <- full_brewer_pal[1:number_categories]  # Take only what we need
  }
  
  # Update the map using maplibre_proxy
  maplibre_proxy("edit_map") |>
    clear_layer("basemap") |>
    add_fill_layer(
      id = "basemap",
      source = full_data(),
      fill_color = step_expr(
        column = "asunum",
        base = "#808080",  # Grey for asunum == 0 (and anything < first break)
        stops = brewer_pal,
        values = asu_breaks,
        na_color = "#444444"
      ),
      fill_opacity = 0.5,
      fill_outline_color = "black",
      tooltip = concat(
        "<strong>ASU Number: </strong>", get_column("asunum"), "<br>",
        "<strong>Tract GEOID: </strong>", get_column("GEOID"), "<br>",
        "<strong>Tract Unemployment Rate: </strong>", number_format(get_column("tract_ASU_urate"), style = "decimal", maximum_fraction_digits = 2), "%<br>",
        "<strong>Tract Unemployment: </strong>", number_format(get_column("tract_ASU_unemp"), style = "decimal", maximum_fraction_digits = 0), "<br>"
      )
    ) |>
    fit_bounds(full_data(), animate = TRUE)
})

```

## Output {data-width=850}

```{r}

```


ASU Review and Finalization
==========================

## Controls {data-width=150}

This screen will allow you to review your selected ASUs.  It uses a threshold of 6.45% unemployment and a population of 10,000, but these numbers can be modified in case they change in the future.

*Generate ASU Summary* will show a table of all ASUs, and whether they pass or fail the ASU definitions for unemployment and population. Please pay attention to these measures - an unemployment rate of 6.449% will display as 6.45%, but will fail the validation.

Note, you *may* go back and forth between the selection and review screens.

When ready to create an output file, the *Generate LSS .txt File* button will create a formatted batch file for upload to LSS.

```{r}

numericInput("asu_ur", "ASU Unemployment Rate Threshold:", value = 6.45, min = 0)

numericInput("asu_pop", "ASU Population Threshold:", value = 10000, min = 0)

actionButton("run_asu_summary", "Generate ASU Summary")
renderUI(HTML("<hr>"))

textOutput("current_save_dir")
renderUI(HTML("<hr>"))

actionButton("lss_txt", "Generate LSS .txt File")
renderUI(HTML("<hr>"))

actionButton("review_txt", "Create Summary CSV")

```

## Output {data-width=850}

```{r}

tableOutput("asu_review")

summarized_asu <- reactiveVal(NULL)

observeEvent(input$run_asu_summary, {
  
  req(full_data())
  
  summary_table <- full_data() %>%
    st_drop_geometry() %>%
    group_by(asunum) %>%
    filter(asunum > 0.5) %>%
    summarize(
      tracts = n(),
      population = sum(tract_pop_cur, na.rm = TRUE),
      lf = sum(tract_ASU_clf, na.rm = TRUE),
      unemp = sum(tract_ASU_unemp, na.rm = TRUE),
      ur = round(unemp/lf*100, 5)
    ) %>%
    ungroup() %>%
    mutate(
      asu_num = round(asunum),
      ur_qualified = if_else(ur >= input$asu_ur, TRUE, FALSE),
      pop_qualified = if_else(population >= input$asu_pop, TRUE, FALSE),
      asu_qualified = ur_qualified & pop_qualified
    )
  
  summarized_asu(summary_table)
})

output$asu_review <- renderTable({
  summarized_asu()
})

observeEvent(input$lss_txt, {
  
  req(full_data())

  # Function to generate text content
  generate_txt <- function() {
    df <- full_data() %>%
      st_drop_geometry() %>%
      select(GEOID, asunum) %>%
      filter(asunum > 0) %>%
      mutate(asunum = as.integer(asunum),
             GEOID = paste0("14000US", GEOID),
             asunum = paste0("SU", state(), sprintf("%04d", asunum)),
             GEOID = paste(asunum, GEOID, sep = " + ")) %>% 
      arrange(asunum, GEOID) %>%
      select(-asunum)

    txt_content <- capture.output(write.table(df, quote=FALSE, row.names=FALSE, col.names = FALSE))
    txt_content <- paste(txt_content, collapse="\n")
    return(txt_content)
  }

  # Get the current save directory from the reactive value
  current_dir <- save_directory()
  
  # Validate directory path
  if (is.null(current_dir) || current_dir == "") {
    showModal(modalDialog(
      title = "No Directory Selected",
      "Please select a save directory first in the Save and Load Data tab.",
      easyClose = TRUE
    ))
    return()
  }

  # Create the content
  txt_content <- generate_txt()

  # Use the selected save directory
  filePath <- file.path(current_dir, "lss_batch_file.txt")
  
  # Ensure directory exists
  if (!dir.exists(current_dir)) {
    tryCatch({
      dir.create(current_dir, recursive = TRUE)
    }, error = function(e) {
      showModal(modalDialog(
        title = "Directory Error",
        paste("Cannot create directory:", current_dir, "\nError:", e$message),
        easyClose = TRUE
      ))
      return()
    })
  }
  
  tryCatch({
    # Write the file
    writeLines(txt_content, filePath)
    
    # Show success message with file location
    showModal(modalDialog(
      title = "File Created Successfully",
      HTML(paste0(
        "<strong>LSS batch file has been saved to:</strong><br>",
        "<code>", filePath, "</code>"
      )),
      easyClose = TRUE
    ))
    
  }, error = function(e) {
    showModal(modalDialog(
      title = "Error Creating File",
      paste("Failed to create LSS file:", e$message),
      easyClose = TRUE
    ))
  })
})

observeEvent(input$review_txt, {

  req(full_data())

  # Function to generate CSV content
  generate_csv <- function() {
    df <- full_data() %>%
      st_drop_geometry() %>%
      arrange(asunum, GEOID)
    
    return(df)
  }
  
  # Get the current save directory from the reactive value
  current_dir <- save_directory()
  
  # Validate directory path
  if (is.null(current_dir) || current_dir == "") {
    showModal(modalDialog(
      title = "No Directory Selected",
      "Please select a save directory first in the Save and Load Data tab.",
      easyClose = TRUE
    ))
    return()
  }

  # Create the content
  df <- generate_csv()

  # Use the selected save directory
  filePath <- file.path(current_dir, "ASU_Review_File.csv")
  
  # Ensure directory exists
  if (!dir.exists(current_dir)) {
    tryCatch({
      dir.create(current_dir, recursive = TRUE)
    }, error = function(e) {
      showModal(modalDialog(
        title = "Directory Error",
        paste("Cannot create directory:", current_dir, "\nError:", e$message),
        easyClose = TRUE
      ))
      return()
    })
  }
  
  tryCatch({
    # Write the CSV file
    write.csv(df, filePath, row.names = FALSE)
    
    # Show success message with file location
    showModal(modalDialog(
      title = "File Created Successfully",
      HTML(paste0(
        "<strong>ASU Review CSV has been saved to:</strong><br>",
        "<code>", filePath, "</code>"
      )),
      easyClose = TRUE
    ))
    
  }, error = function(e) {
    showModal(modalDialog(
      title = "Error Creating File",
      paste("Failed to create CSV file:", e$message),
      easyClose = TRUE
    ))
  })
})


####################
## Future Development

# Export Excel Worksheet with tract details and selections.
# Allow "Save" of data with txt/Excel files so single session is not required.
# Better table formatting.
# highlight high unemployment tracts in non-included areas.
# Check for contiguity of selected ASUs after manual edits.
# Export high-resolution images of ASUs for review.
# Use reactlog package to monitor reactive interactions in the code for cleanup. https://mastering-shiny.org/reactive-graph.html


```



Debugging
=======================================
```{r}
verbatimTextOutput("map_data_preview")
```





